### 2015-03-20 ResponseToReviewers.md

Associate Editor Comments to Author:

Associate Editor
Comments to the Author:
After reading the manuscript and the comments of the two reviewers I think that this approach has high potential but that the work needs some considerable improvements as well. in addition to the comments by the reviewers I suggest to take care of the following:
#- In the beginning, I had some difficulties to follow which different metrics were actually tested against each other in the end and what different information they give us about the patterns and how we could interpret them in ecological terms. There is some text abut this in the beginning of the result section but this is much too late (and to me also an odd place to put this info). I would suggest that this is made clearer early on and maybe it would be good to add a table that lists the indices together with their ecological meaning.

- I moved the paragraph about interpretation to the methods section, and added more explanations, but as I didn't know in advance how Dq will be, I present most interpretations as a discussion. I added a table with the methods used.

# -It should be stated early on in which situations and for which questions this method can be used and it should be highlighted somewhere what assumptions need to be tested to apply it
- The assumptions are stated in methods and I am testing here the power of the test to compare communities, but other possible applications are suggested for Dq to characterize rare species.

### 2015-04-02 ResponseToReviewers.md

# Reviewers comments

Associate Editor Comments to Author:

Associate Editor
Comments to the Author:
After reading the manuscript and the comments of the two reviewers I think that this approach has high potential but that the work needs some considerable improvements as well. in addition to the comments by the reviewers I suggest to take care of the following:
#- In the beginning, I had some difficulties to follow which different metrics were actually tested against each other in the end and what different information they give us about the patterns and how we could interpret them in ecological terms. There is some text abut this in the beginning of the result section but this is much too late (and to me also an odd place to put this info). I would suggest that this is made clearer early on and maybe it would be good to add a table that lists the indices together with their ecological meaning.

* I moved the paragraph about interpretation to the methods section, and added more explanations, but as I didn't know in advance how Dq will be, I present most interpretations as a discussion. I added a table with the methods used.

# -It should be stated early on in which situations and for which questions this method can be used and it should be highlighted somewhere what assumptions need to be tested to apply it
* The assumptions are stated in methods and I am testing here the power of the test to compare communities, but other possible applications are suggested for Dq to characterize rare species.

@@

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Corresponding Author
The author proposes new macroecological metrics – species rank surface (SRS) and generalized dimensions spectrum DqSRS, which is the result of SRS multifractal analysis. New metrics are compared with the classical rank abundance distribution (RAD) and the spectrum of generalized dimensions DqSAD, which reflects the scaling of species abundance distribution (SAD). To demonstrate the possibilities of new metrics, author uses a simple toy model and a more realistic dynamic model of community. Author also has developed a procedure for statistical comparison of RAD, DqSRS and DqSAD for the latter model. This procedure is applied to detect differences between communities with varying degrees of neutrality as well as to demonstrate the differences in power between the metrics.
I highly appreciate the new method of analysis and consider it as having high potential. However, the current manuscript is not suitable for publication. The main drawbacks:

#1. The author offers no meaningful interpretation of dimensions DqSRS, restricting discussion to statements that dimension for large negative q reflect the distribution of rare species, and dimensions for large positive q reflect the distribution of the dominant species. Such lack of interpretation leads to that the new metric is used rather as a "black box", which can or cannot distinguish the communities.
* I added a some more interpretations, and about wath we know and what we don't know about Dq. This is why I added the simple toy models to understand more about how D_q is related to SAD and spatial patterns. pag 7:6_28

#2. Toy examples are chosen poorly. I cannot imagine a community consisting of linear clusters of species ordered by their overall abundance.
* these are examples of very simple species gradients and don't intend to be realistic, and mimic in a basic way an enviromental gradient.

#3. Statistical procedure used for comparison of spectra is very unconventional and raises suspicions in its correctness. For comparison between a pair of two-dimensional curves author uses Anderson-Darling test, which is designed to compare the empirical distribution functions of two samples of one-dimensional random variables, samples must consist of independent observations. From the point of view of Anderson-Darling test the generalized dimensions spectrum Dq is a bimodal distribution with modes at D_-inf and D_+inf and relatively homogeneous distribution between them. Consequently, what is compared by Anderson-Darling test is just a position of the minimum and maximum of spectra. Obviously, this is not a procedure that is needed for spectra comparison. I did not encountered the procedure for the statistical comparison of the generalized dimensions spectra. If the author intends to develop such a procedure, its description and justification must be given with a stronger consideration.
* I added an explanation of why this test can be used with D_q pag 9:10-24

#4. The spectrum of generalized dimensions reflects the spatial scaling of some measure. The use of multifractal analysis implies self-similarity of the studied object, which manifests itself in the power-law form of dependence of partition function Zq on scale. Analysis of the relationship Zq (A) must precede any further analysis. If there is no power-law scaling, the object is not self-similar and its fractal analysis makes no sense. The power law looks like a linear function on a logarithmic scale. The author uses the coefficient of determination R^2 for the analysis of linearity. R^2 is not a measure of linearity, it reflects only the scatter of points around the regression line. The proposed rule of thumb was not justified. 
* I added an explanation about why I use R^2 and delete the rule of thumb. Pag 6:29- Pag 7:1-5

#The author shows figure with Zq(A) scaling only for toy models but not for dynamical models, whereas it is well-studied for neutral models that their species-area relations is triphasic with power law in a narrow range of scales. It is not clear whether there is power-law scaling for dynamical model in studied range of scales.

* I Eliminated figure 4 with R^2 and add figure 6 with fit for neutral model, and also figures 9-13 in the appendix.

@@

I believe that the current version of manuscript requires serious improvement and at this stage must be rejected and resubmitted.

More specific comments and suggestions:

#1. In the description of generalized dimensions spectra it is better to introduce two types (DqSRS and DqSAD) immediately to avoid confusion in formulas 2 and 5. It is better to use notation D1 in formula 3. Formula 5 must specify explicitly that Zq is an average over plots of area A.

* I changed that in the ms. pag 4:25-29 pag 6:1

#2. In the description of SRS details of the algorithm of assigning ranks to equally abundant species must be specified. I think that various options (random assignment of consecutive ranks or assignment of the same  average rank) will have a strong impact on the shape of DqSRS spectrum, especially for negative q. This aspect needs special consideration.

* I added the method used to calculate ranks. pag 6:23-24

#3. In the description of dynamic model (at least in Appendix) the following details must be specified: the boundary conditions, the type of lattice update (synchronous / asynchronous), the algorithm of assignment of competitive ranks in metacommunity. It is also necessary to raise the issue of the model dynamics stationarity, I propose to provide some figures of dynamics of several structural indicators (at least in the appendix).

* More details and figures 1-2 were added in the appendix.

#4. D_0^SRS is not a fractal dimension of SRS (p. 9, l. 20). SRS is a rugged surface and its dimension is more than 2 but less than 3. What is really analyzed is some measure distributed over two-dimensional support. D_0^SRS is a dimension of this support.

* D_0^SRS is the fractal dimension of the species' spatial distribution, which in our model simulations is always 2 because the lattice is always filled. I changed it in the ms. pag 10:23-25

#5. The degree of difference between Dq spectra is not a direct function of Delta rho. It is not correct to analyze relation of power to Delta rho. For example, the great drop in power of DqSAD in Fig. 8 is associated exclusively with the fact that the point Delta rho = 0.9 corresponds to the comparison of simulations rho = 1 vs rho = 0.9, which has very similar spectra.

* I changed figure 8 to a matrix of plots.

#6. The power of methods based on spectra of indices is compared with power of methods based on single components of spectra. It is obvious that the latter will be less powerful. But those methods use different statistical procedures so its direct comparison is not fully correct.

* I added an explanation on why I calculate power for 1 dimension, the idea is to show that single components should not be considered for community comparisons due to its high type I error and low power. When you have different methods to perform a test of a hypothesis you allways prefer the one with more power, so it is commoon to compare power of different methods. pag 10:1-5

#7. Figures 1 and 5 lacks color scheme, making them difficult to read.
* I added the labels and changed color palette

#8. Figure 3 has different scales for DqSAD и DqSRS, making it difficult to compare. As A = epsilon^2 I recommend to use A in both cases.
* I changed x axis to epsilon for both.

#9. RAD is a strictly nonincreasing sequence, but many curves at figure 6 lack this property.
* That was a mistake in calculating averages, I corrected the calculation of ranks.

@@

Reviewer: 2

Comments to the Corresponding Author
Any description of biodiversity that does not explicitly include species abundance distributions (SADs) is  likely to miss important relationships on how species interact with the environment. This is one of the rarest publications in the area of SAD modelling in which author(s) performed an elegant study investigating the scaling of SAD.
The author introduces and innovative framework,  “the species rank surfaces (SRS)”, as a new way to study the scaling of SAD.
Some comments for improvement:

#1)      In the Discussion the author should also highlight the limitations of their modelling process, in particular the fact that it not takes into account: a) asymmetry of the dispersal ability of species; b) the fact that trophic structure and species interactions in communities  may lead to more idiosyncratic processes of community assembly; c) the fact  that the truncated log-normal SAD was not tested.

* This manuscript is about a methodology to compare communities the model was chosen to generate a gradient from  similar and different communities, the comments of the reviewer are very interesting but for a paper about community assembly, that would be another paper...

#2)      The following paper could be cited in line 7 of page 1: Triantis, K.A., Guilhaumon, F. & Whittaker, R.J. (2012). The island species–area relationship: biology and statistics. Journal of Biogeography, 39: 215-23.

#3)      In line 24 of page 1 it seems that one additional reference is missing afer Yakimove et al., 2008)
#4)      In page 14 line 23 “comparable” should be “comparable”
#5)      The author also should revise the English of the text.
#The bibliographic references are a Chaos! The same reference appears in the text in different formats (e.g. Borda-de-Água (2012) and the same correctly as Borda-de-Água et al. (2012). All the bibliographic references should be revised for consistency both in main text and List of Bibliography.
#6)      In the reference list for some strange reason all appear with "et al." in the end. I think that the author should learn to use the tools in Latex (see http://jabref.sourceforge.net/; http://en.wikibooks.org/wiki/LaTeX/Bibliography_Management
#For instance the reference below is a book Chapter:
#Borda-de-Água, L., Hubbell, S.P. & He, F.et al. (2007). Scaling biodiversity under neutrality. Scaling 3 biodiversity, 347–375
#and should be more correctly:
#Borda-de-Água, L., Hubbell, S. P., & He, F. (2007). Scaling biodiversity under neutrality. #In: Storch, D., Marquet, P. A., & Brown. J. H. (eds), Scaling biodiversity. Cambridge University Press, Cambridge, 347-375

* the references were corrected and the ms was read by a native.

### 2015-05-06 ResponseToReviewers2.md

# Reviewers comments resubmited manuscript

Associate Editor Comments to Author:

Associate Editor  
Comments to the Author:
I agree with the two reviewers that the manuscript has greatly improved. Please look at their responses for the few remaining issues concerning the content.
In addition, there are still many language issues and careless mistakes in the manuscript. This needs to be improved and I suggest to find (again) a native speaker for careful proofreading. I mention a few but did not check in detail.
p.4, l.2-3: check grammar
# 4, 26: change to 'first'
# 5, 13: 'graph'
# 6, 29: missing free space
# 7, 9: 'expresses'
# 7, 31: 'took'
# 11, 13: double 'with'
# 15, 21: did not 'make'
# 16, 1: What means: 'low species communities'? 
* communities with low richness

# 16, 2: What means: 'when species are around 100'?
* the number of species is around 100

# 16, 7: 'sizes'

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Corresponding Author

The author has improved the manuscript substantially from the previous version. All my concerns were addressed. The most crucial points are widened discussion of spectra interpretation and clarification of the use of Anderson-Darling criterion for spectra comparison. I believe that AD test is not an ideal solution (for example it cannot distinguish between spectra shifted along each other) but I’m now sure that there will be no inflated type 1 error.

# I think the work now is very close to be accepted but I found several places to be corrected. The most important are plots with scaling of log Zq vs log epsilon in the case of DqSAD (Fig. 3  and 6, left panels). Zq for q < 1 must increase wi th scale and Zq for q > 1 must decrease. Specifically Z0 is just a species richness and so the plot of its scaling is a usual SAR. But on the figures this pattern is reversed. I suspect there was some technical error. Otherwise description of algorithms must be reconsidered.

* For DqSAD equations 4 & 5, we have Zq(A) = Zq(e*e) is the sum of densities of species raised to q in a subplot with side e if q>0 when the area is bigger A'>A the sum Zq(A')>=Zq(A), for q<0 we have Zq(A')<=Zq(A) and this is what we can observe for DqSAD (and also DqSRS) in figures 3 and 6. 
For example if we take only the upper left square to calculate Zq(A):

1 1 2 1
2 1 2 3
2 1 2 3   
1 2 2 1

q=2 e=2  Zq(e*e)= 3*3 + 1
q=2 e=3  Zq(e*e)= 4*4 + 5*5
q=2 e=4  Zq(e*e)= 7*7 + 7*7 + 2*2

Thus Zq is increasing (the exact term is non-decreasing) when q>0

Other minor corrections:

# p. 3, l. 12: change “scaling of Renyi entropies” to “scaling exponents of Renyi entropies”
# p. 3, l. 25: Yakimov et al., 2014 and change reference accordingly
# p. 5, formula 3: I think it is better to specify that this is D1 explicitly (change Dq to D1)
# p. 5, l. 25: change “the abundance” to “the relative abundance”
# p. 6, l. 18: insert “relations” after “species-abundance-area”
# p. 11, l. 5: Fig. 3 shows not an example of fitted Dq but scaling of Zq fitted by linear regressions (and so the figure captions must be corrected)
* I corrected the captions of figures 3 and 6

# p. 14, l. 4: alfa = 0.05 obviously
# p. 14, l. 7: Table 3

Reviewer: 2

Comments to the Corresponding Author
I find the manuscript much improved, the writing more clear and the methods and results much easier to follow. I appreciate the effort the author has made to incorporate the referees criticisms. I make only some minor corrections.
.
Some comments for improvement:
# 1)      Page 12 – line 13 – “with with” should be replaced with “with”
# 2)      Page 6 – line 29 – “(Liebovitch 1998)and” should be replaced with “(Liebovitch 1998) and”
# 3)      the reference below is a book Chapter:
#Borda-de-Água, L., Hubbell, S.P. & He, F.et al. (2007). Scaling biodiversity under neutrality. Scaling 3 biodiversity, 347–375
#and should be more correctly:
#Borda-de-Água, L., Hubbell, S. P., & He, F. (2007). Scaling biodiversity under neutrality, Scaling biodiversity (eds. D. Storch, P.A. Marquet & J.H. Brown) pp. 347-375. Cambridge University Press, Cambridge.

#4)      Dal Bello, M., Maggi, E., Rindi, L., Capocchi, A., Fontanini, D., Sanz-Lazaro, C. & Benedetti-Cecchi, L. 27 (2014). Multifractal spatial distribution of epilithic microphytobenthos on a Mediterranean rocky shore. 28 Oikos, n/a—–n/a
#SHOULD BE:
#Dal Bello, M., Maggi, E., Rindi, L., Capocchi, A., Fontanini, D., Sanz-Lazaro, C. & Benedetti-Cecchi, L. 27 (2014). Multifractal spatial distribution of epilithic microphytobenthos on a Mediterranean rocky shore. Oikos, 124, 477–485.
#5)      In page 21 “Trends in ecology & evolution”, should be “Trends in Ecology & Evolution”
#6)      The reference
#Yakimov, B.N., Iudin, D.I., Solntsev, L.A. & Gelashvili, D.B. (2013). Multifractal analysis of neutral  community spatial structure. Journal of Theoretical Biology.

#Should be
#Yakimov, B.N., Iudin, D.I., Solntsev, L.A. & Gelashvili, D.B. (2013). Multifractal analysis of neutral  community spatial structure. Journal of Theoretical Biology, 343, 44–53.

My best regards to the author,

Paulo A. V. Borges

Azorean Biodiversity Group

cE3c - Centre for Ecology, Evolution and Environmental Changes
University of Azores

Associate Editor Comments to Author:

Associate Editor
Comments to the Author:
After reading the manuscript and the comments of the two reviewers I think that this approach has high potential but that the work needs some considerable improvements as well. in addition to the comments by the reviewers I suggest to take care of the following:
#- In the beginning, I had some difficulties to follow which different metrics were actually tested against each other in the end and what different information they give us about the patterns and how we could interpret them in ecological terms. There is some text abut this in the beginning of the result section but this is much too late (and to me also an odd place to put this info). I would suggest that this is made clearer early on and maybe it would be good to add a table that lists the indices together with their ecological meaning.

* I moved the paragraph about interpretation to the methods section, and added more explanations, but as I didn't know in advance how Dq will be, I present most interpretations as a discussion. I added a table with the methods used.

### 2015-05-10 ResponseToReviewers2.md

Comments to the Corresponding Author

The author has improved the manuscript substantially from the previous version. All my concerns were addressed. The most crucial points are widened discussion of spectra interpretation and clarification of the use of Anderson-Darling criterion for spectra comparison. I believe that AD test is not an ideal solution (for example it cannot distinguish between spectra shifted along each other) but I’m now sure that there will be no inflated type 1 error.

# I think the work now is very close to be accepted but I found several places to be corrected. The most important are plots with scaling of log Zq vs log epsilon in the case of DqSAD (Fig. 3  and 6, left panels). Zq for q < 1 must increase wi th scale and Zq for q > 1 must decrease. Specifically Z0 is just a species richness and so the plot of its scaling is a usual SAR. But on the figures this pattern is reversed. I suspect there was some technical error. Otherwise description of algorithms must be reconsidered.

* I corrected this. The error was produced because I calculated DqSAD=-Tau/(q-1) which is equivalent to DqSAD=Tau/(1-q), thus the slopes were inverted.

### 2015-05-10 Neutral_simulPow.Rmd
correct inverted slopes in Zq vs e

### Plot Neutral Dq Fit
#
setwd("Simul")
zq <-readNeutral_Dq_fit(64,256,500,"L")
plotDqFitG(zq)
setwd(oldcd)
ggsave("DqFit_NeuLog64_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(8,256,500,"L")
plotDqFitG(zq)
setwd(oldcd)
ggsave("DqFit_NeuLog8_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(256,256,500,"L")
plotDqFitG(zq)
setwd(oldcd)
ggsave("DqFit_NeuLog256_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(8,512,500,"L")
plotDqFitG(zq)
setwd(oldcd)
ggsave("DqFit_NeuLog8_512.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(64,512,500,"L")
plotDqFitG(zq)
setwd(oldcd)
ggsave("DqFit_NeuLog64_512.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(256,512,500,"L")
plotDqFitG(zq)
setwd(oldcd)
ggsave("DqFit_NeuLog256_512.png", width=8,height=8,units="in",dpi=600)

>>>>

### Plot Neutral Dq Fit
#
setwd("Simul")
zq <-readNeutral_Dq_fit(64,256,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog64_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(8,256,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog8_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(256,256,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog256_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(8,512,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog8_512.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(64,512,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog64_512.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(256,512,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog256_512.png", width=8,height=8,units="in",dpi=600)
